### 2020-01-01 docs/thesis-project/article.tex
Update article.tex

\subsubsection{Kademlia} \cite{kademlia} is fully decentralized and designed for key-value storage systems, and according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to the least recently seen policy. As a peer receives messages, if the sender is a new acquaintance, that sender is placed on the proper k-bucket. If a replacement is required, then the least seen peer will be evicted, but only if that peer fails to respond to a ping. This policy helps in the defense against attacks where malicious entities flood the network with new dummy peers. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers to learn useful routing information from messages they receive. K-buckets allow routing to be done around failed nodes, providing increased churn resistance. We note that Kademlia does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are when he returns. The standard implementation only removes dated entries when the peer contacts another and obtains no response;

\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired overlay designed for cloud or grid computing architectures. Its foundations are that of the standard Chord, both in terms of identifier generation and the way peers connect, i.e., in ring-like shapes. Self-Chord a few key differences, however, which result from having \textit{ant} agents independently roaming the overlay but collaboratively re-arranging resource location, i.e., data placement is not based solely on the numerical-space relation between data and placement peers. The rearrangement improves lookup operation speed by clustering similar items around neighboring nodes and provides some churn resistance. Other advantages include less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually distribute data appropriately; While redistributing data, Ants also ensure load-balancing, thus improving scalability and robustness since individual peer failures are less likely to lead to massive file loss. According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time. Bio-inspired systems like Self-Chord, diminish or eliminate some of the inherent disadvantages associated with structured or unstructured overlays. In these, simple agents often accomplish complex behavior that would otherwise require time or space consuming algorithms, often with high code complexity. The downside of bio-inspired overlays is that the results of agent actions are not always easy to evaluate or predict. In our proposal, we, too, seek to create a system that employs autonomous, self-healing, and self-organizing behaviors.

@@

\subsubsection{Gia} \cite{gia} seeks to overcome the drawbacks of Gnutella\cite{gnutella-rfc}, which was the first decentralized, unstructured P2P protocol. Gnutella's main issues were the lack of scalability, as resource discovery was made with flooding, and how easily some peers became overloaded when faced with high rates of aggregate queries. Gia proposes the use of super-nodes, which are nowadays also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data. Super-node selection and the construction of the topology around them is dynamic. Gia uses random walks, both to alter the topology as well as perform lookup operations. Random-walks alone are less likely to find appropriate lookup responses unless they are biased to high-degree peers, which in turn can make peer overloading frequent. To solve the issue, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors to help guiding queries in the right direction. The algorithm Gia uses to ensure high-capacity-peers are also high-degree ones depends on a satisfaction function autonomously executed by each peer, which returns a value $s\in\mathclose[0,1\mathclose]$. As long as $s<1$, the peer takes the initiative of calling another random peer, preferring those whose capacity is better than his. During the handshake, any of the intervenients may abort, based on their capacity and the degrees of their neighbors. The callee accepts the caller whenever he does not know enough peers or the caller as more capacity than at least one of his neighbors; to avoid disconnecting poorly connected peers, the callee will replace the highest capacity peer whose capacity is smaller than that of the caller. In order to avoid hot-spots or overloading of peers, a peer is only allowed to send requests to a neighbor if that neighbor has explicitly given him one Token representing one request that neighbor is willing to accept; Each peer allocates tokens at the rate at which they can answer requests. The results show that Gia manages to reduce network overhead, as expected since it does not use flooding, but is also up to five orders of magnitude faster than Gnutella performing lookups. Gia successfully promotes optimal performance and longevity of the system through continuous, peer carried optimizations. However, we identify one fundamental problem with Gia, related to the systems attempt to load-balance super-peers and implicit free-rider avoidance\footnote{Peers who consume more community resources than they contribute with.}; According to their capacity, peers receive tokens from their neighbors, hence decreasing the number of tokens received by low-capacity peers and, thus, system usability.

\subsubsection{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols. Like in Self-Chord, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded. The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals. BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed. The overlay management can be viewed in three major components.

The first component are the peers themselves. Each peer keeps two structures, a   fixed-size $\alpha$-table, i.e., a partial view of the overlay retaining neighborhood information used to evaluate the redundant connections or the need for new ones, and, a neighborhood-set \textit{N}, containing peer identifiers. Peers contribute to the optimization by rearranging local connections according to the connection rule, which reduces the diameter of the overlay, and disconnection rule, which discards redundant connections.

The second component are the ant agents, which have multiple species; \textit{discovery ants} are randomly spawned by peers and live for a limited time. They wander across the overlay, collecting information about its topology and update the $\alpha$-tables of peers they meet along the way. \textit{Construction ants} act as bootstrappers to the system. When a new peer \textit{i} wants to connect to a peer \textit{j}, he sends one of these ants, if \textit{j} cannot establish the connection because he would violate a degree constraint, he forwards the ant to his lowest-degree neighbor. When a \textit{j'} accepts the ant, he sends it back to \textit{i}, and the procedure complete, i.e., they are now on each others' neighborhood-sets. \textit{Optimization ants} establish connections according to the connection rule. In this case, a peer \textit{i} wanting to connect to peer \textit{j} sends him this ant, and \textit{j} only accepts or refuses the ant. If he accepts, he sends the ant back to \textit{i}. \textit{Unlink ants} remove existing connections between peers because of the disconnection rule applies or because a peer wants to leave the overlay. When this ant arrives at its destination, it removes all sender's information from the $\alpha$-table and the neighborhood-set. \textit{Update Neighbors Ants} are spawned by a peer whenever he establishes or tears-down a connection with some other. These ants visit all of that peer's neighbors and update his information on their respective $\alpha$-table. Finally, \textit{Ping Ants} are periodically exchanged between peers to keep their connections alive in low traffic situations.

The last component are pheromone trails, which evaporate overtime. A trail is a value assigned to a connection between peers, whenever an ant agent, or a query, walks over that connection, they increase the pheromone concentration increases on both ends. Discovery ants will tend to follow paths with less pheromone concentration, thus incentivizing full network coverage during the exploration. When a peer detects that for a given neighbor, its pheromone concentration has wholly evaporated, that neighbor is assumed to have left the network without warning. Eventually, each peer who neighbored the disconnected peer will independently initiate a recovery protocol by sending construction ants to the neighbors of the disconnected peer, reorganizing the overlay, and avoiding network partitioning. Thus pheromone trails provide seamless error resolution.

BlatAnt proved to accomplish its goals, quickly converging to stable overlays with bounded-diameter, even under very dynamic conditions, resulting in faster lookups due to the reduction of flood messages but also due to caching. Note that BlatAnt can leverage random walks such as the ones used in Gia for even less network overhead. BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and because caching is out of the scope of our proposal.\newline

>>>>

\subsubsection{Kademlia}\cite{kademlia} is fully decentralized and designed for key-value storage systems, and according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to the least recently seen policy. As a peer receives messages, if the sender is a new acquaintance, that sender is placed on the proper k-bucket. If a replacement is required, then the least seen peer will be evicted, but only if that peer fails to respond to a ping. This policy helps in the defense against attacks where malicious entities flood the network with new dummy peers. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers to learn useful routing information from messages they receive. K-buckets allow routing to be done around failed nodes, providing increased churn resistance. We note that Kademlia does not implement an effective mechanism to keep dated entries out of a peer's k-buckets, meaning that the longer a peer stays offline, the worse is k-buckets are when he returns. The standard implementation only removes dated entries when the peer contacts another and obtains no response;

\subsubsection{Self-Chord}\cite{selfchord} is a bio-inspired overlay designed for cloud or grid computing architectures. Its foundations are that of the standard Chord, both in terms of identifier generation and the way peers connect, i.e., in ring-like shapes. Self-Chord a few key differences, however, which result from having \textit{ant} agents independently roaming the overlay but collaboratively re-arranging resource location, i.e., data placement is not based solely on the numerical-space relation between data and placement peers. The rearrangement improves lookup operation speed by clustering similar items around neighboring nodes and provides some churn resistance. Other advantages include less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually distribute data appropriately; While redistributing data, Ants also ensure load-balancing, thus improving scalability and robustness since individual peer failures are less likely to lead to massive file loss. According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time. Bio-inspired systems like Self-Chord, diminish or eliminate some of the inherent disadvantages associated with structured or unstructured overlays. In these, simple agents often accomplish complex behavior that would otherwise require time or space consuming algorithms, often with high code complexity. The downside of bio-inspired overlays is that the results of agent actions are not always easy to evaluate or predict. In our proposal, we, too, seek to create a system that employs autonomous, self-healing, and self-organizing behaviors.

@@

\subsubsection{Gia}\cite{gia} seeks to overcome the drawbacks of Gnutella\cite{gnutella-rfc}, which was the first decentralized, unstructured P2P protocol. Gnutella's main issues were the lack of scalability, as resource discovery was made with flooding, and how easily some peers became overloaded when faced with high rates of aggregate queries. Gia proposes the use of super-nodes, which are nowadays also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data. Super-node selection and the construction of the topology around them is dynamic. Gia uses random walks, both to alter the topology as well as perform lookup operations. Random-walks alone are less likely to find appropriate lookup responses unless they are biased to high-degree peers, which in turn can make peer overloading frequent. To solve the issue, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors to help guiding queries in the right direction. The algorithm Gia uses to ensure high-capacity-peers are also high-degree ones depends on a satisfaction function autonomously executed by each peer, which returns a value $s\in\mathclose[0,1\mathclose]$. As long as $s<1$, the peer takes the initiative of calling another random peer, preferring those whose capacity is better than his. During the handshake, any of the intervenients may abort, based on their capacity and the degrees of their neighbors. The callee accepts the caller whenever he does not know enough peers or the caller as more capacity than at least one of his neighbors; to avoid disconnecting poorly connected peers, the callee will replace the highest capacity peer whose capacity is smaller than that of the caller. In order to avoid hot-spots or overloading of peers, a peer is only allowed to send requests to a neighbor if that neighbor has explicitly given him one Token representing one request that neighbor is willing to accept; Each peer allocates tokens at the rate at which they can answer requests. The results show that Gia manages to reduce network overhead, as expected since it does not use flooding, but is also up to five orders of magnitude faster than Gnutella performing lookups. Gia successfully promotes optimal performance and longevity of the system through continuous, peer carried optimizations. However, we identify one fundamental problem with Gia, related to the systems attempt to load-balance super-peers and implicit free-rider avoidance\footnote{Peers who consume more community resources than they contribute with.}; According to their capacity, peers receive tokens from their neighbors, hence decreasing the number of tokens received by low-capacity peers and, thus, system usability.

\subsubsection{BlatAnt}\cite{blatant} is a bio-inspired overlay designed for grid computing architectures where resource discovery needs to be efficient; BlatAnt focuses solely on network bounded-diameter optimization as means of minimizing network overhead and lookup latency, using swarm intelligence to support flood-like discovery protocols. Like in Self-Chord, ant agents optimize and maintain the overlay; conversely, the ants do not change the location of data items, but instead reorganize peer connections in such a way that the global overlay uses as little of them as possible. The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller. BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches, as an alternative to semantic clustering such as the ones used in Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed. The overlay management can be viewed in three major components. The first component are the \textit{peers} themselves. Each peer keeps two structures, a fixed-size $\alpha$-table, i.e., a partial view of the overlay retaining neighborhood information used to evaluate the redundant connections or the need for new ones, and, a neighborhood-set \textit{N}, containing peer identifiers. Peers contribute to the optimization by rearranging local connections according to the connection rule, which reduces the diameter of the overlay, and disconnection rule, which discards redundant connections. The second component are the ants, which have multiple species; \textit{discovery ants} are randomly spawned by peers and live for a limited time. They wander across the overlay, collecting information about its topology and update the $\alpha$-tables of peers they meet along the way. \textit{Construction ants} act as bootstrappers to the system. When a new peer \textit{i} wants to connect to a peer \textit{j}, he sends one of these ants, if \textit{j} cannot establish the connection because he would violate a degree constraint, he forwards the ant to his lowest-degree neighbor. When a \textit{j'} accepts the ant, he sends it back to \textit{i}, and the procedure complete, i.e., they are now on each others' neighborhood-sets. \textit{Optimization ants} establish connections according to the connection rule. In this case, a peer \textit{i} wanting to connect to peer \textit{j} sends him this ant, and \textit{j} only accepts or refuses the ant. If he accepts, he sends the ant back to \textit{i}. \textit{Unlink ants} remove existing connections between peers because of the disconnection rule applies or because a peer wants to leave the overlay. When this ant arrives at its destination, it removes all sender's information from the $\alpha$-table and the neighborhood-set. \textit{Update Neighbors Ants} are spawned by a peer whenever he establishes or tears-down a connection with some other. These ants visit all of that peer's neighbors and update his information on their respective $\alpha$-table. Finally, \textit{Ping Ants} are periodically exchanged between peers to keep their connections alive in low traffic situations. The last component are \textit{pheromone trails}, which evaporate overtime. A trail is a value assigned to a connection between peers, whenever an ant or query, walks over that connection, they increase the pheromone concentration on both ends. Discovery ants will tend to follow paths with less pheromone concentration, thus incentivizing full network coverage during the exploration. When a peer detects that for a given neighbor, its pheromone concentration has wholly evaporated, that neighbor is assumed to have left the network without warning. Eventually, each peer who neighbored the disconnected peer will independently initiate a recovery protocol by sending construction ants to the neighbors of the disconnected peer, reorganizing the overlay, and avoiding network partitioning. Thus pheromone trails provide seamless error resolution. BlatAnt quickly converges to stable overlays with bounded-diameter, even in very dynamic conditions, resulting in faster lookups due to a decrease in flood messages and also due to caching. Note that BlatAnt could also leverage random walks such as the ones used in Gia for even less network overhead. BlatAnt also introduces optimizations over caching procedures; we do not review them here, because they are based on complex profile similarities and caching is out of the scope of our proposal.
